\vskip-0.6in Trade Agreements in the Shadow of Lobbying\protect\footnote{Formerly circulated under the title ``Trade Agreements, Lobbying and Separation of Powers''}

### 2016-02-01 draft.tex
add paragraph on applicability to parliamentary regimes #10

Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the binary decision by the legislature about whether to abide by or break the trade agreement is modeled on the ``Fast Track Authority'' that the U.S. Congress granted to the Executive branch almost continuously from 1974-1994 and then again as ``Trade Promotion Authority'' from 2002-2007.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries in which the formation and maintenance of trade policy requires ratification by a politically susceptible body either at home or in a trading partner.

>>>>

Although the political process here most closely matches that of the United States in the post-war era,\footnote{In particular, the binary decision by the legislature about whether to abide by or break the trade agreement is modeled on the ``Fast Track Authority'' that the U.S. Congress granted to the Executive branch almost continuously from 1974-1994 and then again as ``Trade Promotion Authority'' from 2002-2007.} the model is broadly applicable because the central dynamic that emerges---that trade agreements are formed in the \textit{shadow} of lobbying pressure instead of in direct response to it---appears to be present in a wide range of countries in which the formation and maintenance of trade policy requires \textit{approval} by a politically susceptible body either at home or in a trading partner.

The results apply beyond separation-of-powers presidential systems because the presence of veto points is what is essential for the dynamic posited here to occur. Although the number of formal veto points tends to be lower in, for instance, parliamentary systems, veto points are typically not absent altogether (Henisz and Mansfield 2006). In reality, constraints on the executive are almost always present in democratic regimes, although they may operate in different ways, e.g. over questions of government dissolution and formation in parliamentary regimes. It is important to consider that constraints may be formal or informal, and they may come from legislative, regulatory, judicial, sub-national or even other executive bodies.
